Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2016-10-28 manuscript.tex
latest

The stress archetype is more ambiguous, and has been the subject of an extensive debate in the plant ecology literature (the ``Grime-Tilman'' debate \citep{aerts_1999}). In Grime's view, stressful environments impose such severe challenges that they are only barely survivable.

In our model, this can be expressed by $b_i\ll 1$ and $b_i/d_i\approx 1$. The propagule density $L$ (as well as $N/K$) are suppressed to such low levels that there is essentially no competition between individuals. The severity of the stress is such that mutations which appreciably improve $b_i$ are extremely rare, so $b_i$ is constrained to remain low.

The alternative view is that the stress archetype should rather be interpreted as a reduction in the maximum number of individuals that can be supported by the environment \citep{taylor_1990}; in our model, this means a greater per-individual territorial requirement represented by a lower $K$ for a given amount of space. The stress implies that $b$ is low

For example, in the commonly cited case that the stress is induced by a scarcity of consumable resources, it is argued that competition for resources would likely be intense, and the stressed population should actually be regarded as having a high population density \citep{taylor_1990}.

That is, $N$ under stress is much lower than  under ideal conditions, but it is not much lower than $K$ for the stressful environment. Since our model accommodates both of these alternatives, we include them both here.

>>>>

The stress archetype is more ambiguous, and has been the subject of an extensive debate in the plant ecology literature (the ``Grime-Tilman'' debate; \citealt{aerts_1999}). Stressful environments severely restrict growth and reproduction, so that $b_i\ll 1$ \cite{grime_1974,grime_1977}. Mutations which appreciably improve $b_i$ will be either non-existent or extremely unlikely, so $b_i$ is constrained to remain low. In Grime's view, under these conditions the rate at which propagules successfully develop to adulthood cannot appreciably exceed the mortality rate. This implies $b_i/d_i\approx 1$ in our model, and so the propagule density $L$ is suppressed to such low levels that there are essentially no territorial contests occurring.

The alternative view is that stressful environments simply have a lower carrying capacity \citep{taylor_1990}; in our model, this means a greater per-individual territorial requirement represented by a lower $K$ for a given amount of space. In particular, it is argued that when stress is induced by a scarcity of consumable resources, competition for those resources would likely be intense. Thus, $b_i$ need not be particularly close to $d_i$, but contests for consumable resources at the juvenile phase would kill off most propagules before adulthood. In other words, the stressed population is a high density population where competition is important \citep{taylor_1990} (that is, high density relative to $K$, not relative to ideal conditions). We include both of these alternatives in our discussion here.
